---
title: "2_beta_diversity"
author: "Pedro Beschoren da Costa"
date: "August 25, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

#General settings
##load libraries

[MA]: Load only packages that you'll need in this script
```{r}
#Set working directory to project directory 
setwd("./")
getwd() #ok

#Handling data 
library("dplyr")
library("tibble")
library("stringr")  # to wrangle string vectors
library("tidyr")
library("ggrepel")  # to avoid legends overlapping in your plot
library("ggpubr")
library("purrr")  # has map() to select table elements

#Microbiome
library("vegan")  # for several essential statistical tests
library("forcats")
library("metagMisc")  # lets you create lists of split phyloseq objects
library("phyloseq")
library("decontam")
library("metagenomeSeq")
library("Biostrings")
library("mikropml")
library(EcolUtils)

#Visuals
library("ggplot2")

#stats
library("agricolae")  # includes some ANOVA post-hoc options
library("minpack.lm")  # lets you do some HSD tests, output is a nice table
library("car")  # for Levene's test

```

## Load data
[MA]: Now, get the data from script 1, only the ps list that you need 
```{r}

#Load 'normalization_listed', a ps list with CSS, rarefied and non-rarefied
normalization_listed <- readRDS("./R output/01_phyloseq_objects/normalization_listed.rds")

#Check ps objects
normalization_listed$physeq_filtered #original ps, non-normalized, full dataset, same as physeq_filtered
normalization_listed$physeq_filtered_rarefied #rarefied, full dataset
normalization_listed$physeq_filtered_CSS #CSS-normalized, full dataset

```

Let's get the summary of reads in the raw phyloseq
```{r}

# Add new variable 'Sp' as a short name for species
normalization_listed <- lapply(normalization_listed, function(ps){
                        # Add new column 'Sp' with short species names
                        sample_data(ps)$Sp <- ifelse(sample_data(ps)$Plant_species == "Arabidopsis_thaliana", "At",
                                          ifelse(sample_data(ps)$Plant_species == "Brassica_oleraceae", "Bo", NA))

                        # Check the first few rows
                        print(head(sample_data(ps)$Sp))
                        
                        #return
                        return(ps)})

#Now add other variable ('Sp_Comp') with the combination of Sp and root compartment (Sample_type)
normalization_listed <- lapply(normalization_listed, function(ps){
                        # Add new column 'Sp' with short species names
                        sample_data(ps)$Sp_Comp <- paste(sample_data(ps)$Sp,
                                              sample_data(ps)$Sample_type, 
                                              sep = "_")

                        # Check the first few rows
                        print(head(sample_data(ps)$Sp_Comp))
                        
                        #return
                        return(ps)})

#Add as well the number of reads and the number of ASVs 
normalization_listed <- lapply(normalization_listed, function(ps){
                        # Add new column 'library_size' 
                        sample_data(ps)$library_size <- sample_sums(ps)
                        
                        #Add new column 'Num_ASVs'
                        sample_data(ps)$Num_ASV <- apply(otu_table(ps), 2, function(x) sum(x > 0))

                        # Check the first few rows
                        print(head(sample_data(ps)))
                        
                        #return
                        return(ps)})

#let's check sample_data
lapply(normalization_listed, function(ps){head(sample_data(ps))})
```

[MA]: Let's split it by Plant Species and Root compartment. 
In the metadata 'Root' is Endosphere and 'Soil' is actually Rhizosphere. 
```{r}

#split it by Species and Root community to get number of ASVs per sub-community
ps_root_soil <- phyloseq_sep_variable(normalization_listed$physeq_filtered, variable = "Sp_Comp")

#get summary
#make a df 
sample_data <- normalization_listed$physeq_filtered %>% 
    sample_data() %>% 
    data.frame()

#let's quickly check how many replicates we have left per Sp and Root compartment (Root=Endosphere, Soil=Rhizosphere)
#and how is the library size amongst them. [part of Supplementary Table S1]
sample_data %>% 
    group_by(Sp, Sample_type)%>%
    dplyr::summarise(Replicates = n(),
              mean_library_size = mean(library_size),
              stdev_libsize =sd(library_size))

```

## Settings for plots 
[MA]: Load general settings for plots (axis, treatment colors, etc)
```{r}

#Set plots theme
ggplot2::theme_set(theme_bw())

# Set default axis looks
axis_looks <- theme(axis.text.x = element_text(colour = "black", size = 9,
                                               face = "bold", angle=0, hjust=0.5))+ #text in X
  theme(axis.text.y = element_text(colour = "black", size = 9, face = "bold"))+ #text in y
  theme(axis.title=element_text(size=11, face = "bold"))+ #all texts
  theme(axis.title.y = element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0)))+
  theme(panel.grid.major = element_blank(), #no grid
        panel.grid.minor = element_blank()) #no grid

#Set colors
color_treatments <- c("#d7d7d7", "#f2c476", "#bd5a28", "#7aaf91") #"Control", "MeJA 0.1mM", "MeJA 1mM", "P. brassicae OS"
color_species <- c("#006c3a", "#309190") #at & bo, respectively
color_compartment <- c("#B28F45", "#846475") #endo & rhizo, respectively (not really used)

```


# 2.0 - basic beta diversity analysis
On this script we will evaluate beta diversity - the differences in the microbial community compositions across samples. we will make ordination plots and multivariate tests


## 2.1 - Beta Diversity plots
Beta diversity plots are the beating heart or microbiome analysis. here you will be able to visually tell if communities differ according to treatment or not. It can be a very long topic but here I only use one option.

```{r}

# Let's make a Non-Metric Multidimensional Scaling (NMDS) of all our samples based on CSS normalization
nmds_rootAndSoil <- phyloseq::ordinate(normalization_listed$physeq_filtered_CSS,
  method = "NMDS", # this method has few assumptions and readly accepts different data structures
  distance = "bray", # bray-curtis distance is suitable for sparse data - such as the zero-inflated microbiome data we have
  try = 200, # number of iterations
  autotransform = TRUE
) # automatically transforms your data, if needed. reduces weight of outliers
# weakties = FALSE prevests tress from colapsing into zero

#First get a short version of sample data df to be used as metadata for NMDS
df_info <- sample_data(normalization_listed$physeq_filtered_CSS) %>% 
               data.frame() %>% 
               dplyr::select(Plant_species, Sample_type, Sp_Comp, MeJA_treatment)%>%
               mutate(sample_name=rownames(.))

#[MA] call function to get coordenates 
get_NMDS_coordenates <- function(x,y){
  
  #x is a phyloseq
  #should be a string in between "" with the distance (i.e. "bray")
  
  df_nmds <- lapply(y, function(dist){
    ord <- ordinate(x, "NMDS", distance=dist, try=200, autotransform=TRUE)
    nMDS <- plot_ordination(x, ord, color="MeJA_treatment")
    df_NMDS <- nMDS$data
    df_NMDS <- df_NMDS %>% dplyr::select(Plant_species, Sample_type, Sp_Comp,MeJA_treatment,NMDS1, NMDS2)
    df_NMDS$sample_name <- rownames(df_NMDS)
    df_NMDS <- df_NMDS %>% relocate(sample_name)
    return(df_NMDS)
    })

  #ord will be a list 
  return(df_nmds)
}


#[MA] get coordinates
df_NMDS <- get_NMDS_coordenates(normalization_listed$physeq_filtered_CSS, "bray")
df_NMDS <- df_NMDS[[1]] #it was a list before idk why
str(df_NMDS)

df_NMDS$Plant_species <- as.factor(df_NMDS$Plant_species) #change it to factor

#Check order
df_NMDS$Sample_type #Order: Endo and Rhizo
df_NMDS$Plant_species #Order: At and Bo

#plot NMDS 
ggplot(df_NMDS, aes(x=NMDS1,y=NMDS2, 
                    fill=MeJA_treatment, #inside
                    shape=Plant_species,
                    color=Sample_type)) + #outline for root compartment
  geom_point(aes(color=Sample_type, fill=MeJA_treatment), size=5)+ # individual sample values
  scale_shape_manual(values=c(21,24))+ #cirle At, Triangle Bo
  scale_color_manual(values = c("white", "black"))+ #white is endo, black rhizo
  scale_fill_manual(values=color_treatments)+ #inside
  ggtitle("NMDS_full")+
  axis_looks+
  theme(legend.position = "none")
  

#[MA]save plot; Figure 1B
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/02_beta_diversity",
  scale = 1.5,
  width = 1000,
  height = 900,
  units = "px",
  dpi = 300),
x=c("NMDS_Full.svg","NMDS_Full_16S.png"))

```


### 2.1.1 - custom NMDS function for automation of split datasets

```{r}

#Now get a NMDS for each sub-community, for this make a list of 4 ps
ps_list_CSS2 <- phyloseq_sep_variable(normalization_listed$physeq_filtered_CSS, variable = "Sp_Comp")

#call function again to change metadata and only select MeJA treatment
get_NMDS_coordenates <- function(x,y){
  
  #x is a phyloseq
  #should be a string in between "" with the distance (i.e. "bray")
  
  df_nmds <- lapply(y, function(dist){
    ord <- ordinate(x, "NMDS", distance=dist, try=200, autotransform=TRUE)
    nMDS <- plot_ordination(x, ord, color="MeJA_treatment")
    df_NMDS <- nMDS$data
    df_NMDS <- df_NMDS %>% dplyr::select(MeJA_treatment,NMDS1, NMDS2)
    df_NMDS$sample_name <- rownames(df_NMDS)
    df_NMDS <- df_NMDS %>% relocate(sample_name)
    return(df_NMDS)
    })

  #ord will be a list 
  return(df_nmds)
}

#[MA] get coordinates with lapply
df_NMDS_split <- lapply(ps_list_CSS2, function(ps){
          df <- get_NMDS_coordenates(ps, "bray")
          df <- df[[1]]
          return(df)})

#[MA] Now, get centroids 
centroids <- lapply(df_NMDS_split, function(df){
  df <- aggregate(cbind(NMDS1, NMDS2) ~ MeJA_treatment, data=df, FUN=mean)
  return(df)})

#[MA] Finally, get full df to be able to draw segments for this use mapply to go over two lists
df_NMDS_split_full <- mapply(function(df_NMDS_split, centroid_df) {
  merge(df_NMDS_split, centroid_df,
        by = "MeJA_treatment", all.x = TRUE, suffixes=c("", "_centroid"))},
  df_NMDS_split, centroids, SIMPLIFY = FALSE)

df_NMDS_split_full <- lapply(df_NMDS_split_full, function(list){ as.data.frame(list)})
  

#Now, plot the NMDS, one per Sp-Root
AT_Endo <- ggplot(df_NMDS_split_full$At_Root, aes(x=NMDS1,y=NMDS2, 
                    fill=MeJA_treatment,
                    color=MeJA_treatment)) + #outline
  geom_segment(data = df_NMDS_split_full$At_Root,
               mapping = aes(x=NMDS1, y=NMDS2, 
               xend = NMDS1_centroid, yend = NMDS2_centroid), linewidth=0.8, alpha=0.9)+
  geom_point(aes(fill=MeJA_treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.95)+ # individual sample values
  geom_point(aes(x=NMDS1_centroid, y = NMDS2_centroid),
             size = 5, alpha=0.9, shape=21, color="black",stroke = 0.4,) + # centroids
  scale_fill_manual(values = color_treatments)+ #outline
  scale_color_manual(values = color_treatments)+ #outline
  ggtitle("AT_Endo")+
  axis_looks+
  theme(legend.position = "none")

AT_Endo


BO_Endo <- ggplot(df_NMDS_split_full$Bo_Root, aes(x=NMDS1,y=NMDS2, 
                    fill=MeJA_treatment,
                    color=MeJA_treatment)) + #outline
  geom_segment(data = df_NMDS_split_full$Bo_Root,
               mapping = aes(x=NMDS1, y=NMDS2, 
               xend = NMDS1_centroid, yend = NMDS2_centroid), linewidth=0.8, alpha=0.9)+
  geom_point(aes(fill=MeJA_treatment), shape=24, color="black", stroke = 0.3,
             size=3, alpha=0.95)+ # individual sample values
  geom_point(aes(x=NMDS1_centroid, y = NMDS2_centroid),
             size = 5, alpha=0.9, shape=24, color="black",stroke = 0.4,) + # centroids
  scale_fill_manual(values = color_treatments)+ #outline
  scale_color_manual(values = color_treatments)+ #outline
  ggtitle("BO_Endo")+
  axis_looks+
  theme(legend.position = "none")

BO_Endo 


AT_Rhizo <- ggplot(df_NMDS_split_full$At_Soil, aes(x=NMDS1,y=NMDS2, 
                    fill=MeJA_treatment,
                    color=MeJA_treatment)) + #outline
  geom_segment(data = df_NMDS_split_full$At_Soil,
               mapping = aes(x=NMDS1, y=NMDS2, 
               xend = NMDS1_centroid, yend = NMDS2_centroid), linewidth=0.8, alpha=0.9)+
  geom_point(aes(fill=MeJA_treatment), shape=21, color="black", stroke = 0.3,
             size=3, alpha=0.95)+ # individual sample values
  geom_point(aes(x=NMDS1_centroid, y = NMDS2_centroid),
             size = 5, alpha=0.9, shape=21, color="black",stroke = 0.4,) + # centroids
  scale_fill_manual(values = color_treatments)+ #outline
  scale_color_manual(values = color_treatments)+ #outline
  ggtitle("AT_Rhizo")+
  axis_looks+
  theme(legend.position = "none")

AT_Rhizo


BO_Rhizo <- ggplot(df_NMDS_split_full$Bo_Soil, aes(x=NMDS1,y=NMDS2, 
                    fill=MeJA_treatment,
                    color=MeJA_treatment)) + #outline
  geom_segment(data = df_NMDS_split_full$Bo_Soil,
               mapping = aes(x=NMDS1, y=NMDS2, 
               xend = NMDS1_centroid, yend = NMDS2_centroid), linewidth=0.8, alpha=0.9)+
  geom_point(aes(fill=MeJA_treatment), shape=24, color="black", stroke = 0.3,
             size=3, alpha=0.95)+ # individual sample values
  geom_point(aes(x=NMDS1_centroid, y = NMDS2_centroid),
             size = 5, alpha=0.9, shape=24, color="black",stroke = 0.4,) + # centroids
  scale_fill_manual(values = color_treatments)+ #outline
  scale_color_manual(values = color_treatments)+ #outline
  ggtitle("BO_Rhizo")+
  axis_looks+
  theme(legend.position = "none")

BO_Rhizo 


#[MA] Make it a single figure 
ggarrange(AT_Endo, BO_Endo, AT_Rhizo, BO_Rhizo, ncol=4)

#[MA] Save plot
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/02_beta_diversity",
  scale = 1.8,
  width = 2000,
  height = 500,
  units = "px",
  dpi = 300),
x=c("NMDS_SpComp.svg","NMDS_SpComp.png"))

```


## 2.2 - Beta Dispersion

Beta dispersion indicates variance in the multidimensional space. It essentially measures the spread of your data points in the ordination, according your different treatments. As you might expect, statistical tests will have a hard time comparing a treatment with high beta dispersion (wide, sparse datapoint cloud) against a treatment with low beta dispersion (clustered, compact datapoint cloud), just like it is difficult to compare large error bars to small error bars.  You can test for homogeniety of this variance, similarly to a levene test of homogeniety before an anova. 

What if you don't have homogeniety of variances? then your permanova results are not as reliale. This can be difficult to solve. there are a few transformations you could try (log or sqrt of your counts), outliers you could remove (specially if you observed that a particular outlier was specially unhealthy) or just split your permanovas accordingly (if soil samples have higher beta dispersion than root samples, test root and soil in separate permanovas). sometimes however these alternatives do no solve the problem - proceed with your tests as normal but be cautious with your conclusions.

with this chunk we evaluate Beta Dipersion Metrics & plots 
```{r}

# this will test beta dispersions over a list by considering a factor
#### -------- soft-coded, 3 lists of phyloseq objects ---###
beta_disp_plotAndTest <- function(phyloseq_list, group) {
  # phyloseq_list = a list of phyloseq objects
  # group = the variale you want to test the beta dispersion of, in quotes
  beta_disper_list <- lapply(phyloseq_list, function(x) {
    betadisper(phyloseq::distance(t(otu_table(x)), method = "bray"), sample_data(x)[[group]])
  }) # selects only column "group""

  # gets the names of the list
  tiles_list <- names(beta_disper_list)

  # runs anova on beta dispersions
  get_p <- lapply(beta_disper_list, function(x) {
    anova(x, permutations = 999)
  })


  p_dispersion <- map(get_p, 5) # gets the p value of the dispersion test
  p_dispersion <- p_dispersion[!is.na(p_dispersion)] # removes the generated NA


  # runs anova on beta dispersions
  bet_disp_PCOa_plot <- mapply(function(x, y, z) {
    plot(x,
      main = y,
      sub = z,
      xlab = "p value for homogeniety test:",
      ylab = "PCoA"
    )
  },
  x = beta_disper_list,
  y = tiles_list,
  z = p_dispersion,
  SIMPLIFY = FALSE
  )

  # runs anova on beta dispersions
  bet_disp_boxplot <- mapply(function(x, y) {
    boxplot(x, main = y)
  },
  x = beta_disper_list,
  y = tiles_list,
  SIMPLIFY = FALSE
  )


  return(list(bet_disp_PCOa_plot, bet_disp_boxplot))
}

```



```{r}

#[MA] As PERMANOVA will be donde in CSS-normalized data, split the phyloseq of physeq_filtered_CSS by Species and Root community #In old code this was the same as ps_list_CSS
#Now changed in this way as phyloseq_sep_variable doesn't accept anymore to split by two variables, now it has to be 1
#therefore changed from phyloseq_sep_variable(ps, variable = c("Plant_species", "Sample_type")) to 
#phyloseq_sep_variable(ps, variable = "Sp_Comp"), which is exactly the same 

ps_list_CSS <- phyloseq_sep_variable(normalization_listed$physeq_filtered_CSS, variable = "Sp_Comp")


# now that we have a custom function, we can run it across all lists and variables
set.seed(5235)
dip_result3 <- beta_disp_plotAndTest(ps_list_CSS, "MeJA_treatment")

#check
dip_result3 #all passed

```


## 2.3 - Run PERMANOVA 

A permutation anova will tell if the differences in the microbial community structure are significant or not. they will essentially help you separate the data clouds of your ordination with confidence levels


you will need to run, test and check several different models and data slices to have final insight into the dataset you are evaluating. get used with testing multiple models!

```{r}

#[MA] Check it in the non-split ps that is normalized by CSS 
physeq_filtered_CSS <- normalization_listed$physeq_filtered_CSS

# Running the permanova with vegan::adonis2() on a single phyloseq object is very simple
metadata <- as(sample_data(physeq_filtered_CSS), "data.frame")

# let's first define permitation design with permute::how()
perm2 <- how(
  within = Within(type = "free"), # permute observations freely within the blocks. got to ask Fred how to handle grid arrangements!
  #blocks = metadata$Block, # define blocks, silenced as it doesn't change the output
  nperm = 999
) # n of permutations
set.seed(5235)

# now let's calculate the permanova in the full ps 
three_way_permanova <- adonis2(phyloseq::distance(t(otu_table(physeq_filtered_CSS)), method = "bray") # this is your distance matrix of OTU abundances
~ MeJA_treatment * Plant_species * Sample_type, # this is your model
permutations = perm2, # we jsut defined the permutation settings above, you place them here
data = metadata
) # metadata for the community matrix

#check
three_way_permanova #[This is Table S2]


#### -------- soft-coded, 3 lists of phyloseq objects with 1 formula ---###

# let's check now if sample type differ per species
permanova_with_list <- function(phyloseq_list, rhs_model) {
  # phyloseq_list = list of phyloseq objects
  # RHS_model = right hand side model, such as MeJA_treatment*Sample_type
  lapply(phyloseq_list, function(x) {
    lhs_data <- phyloseq::distance(t(otu_table(x)), method = "bray")
    rhs_model_in <- paste(rhs_model)
    form <- as.formula(paste("lhs_data~", paste(rhs_model_in))) # getting the formulat properly evaluated as custom string is tricky
    output <- adonis2(form,
      data = as(sample_data(x), "data.frame"), # changing with as.data.frame is insufficient
      permutations = how(within = Within(type = "free"),
                         blocks = as(sample_data(x), "data.frame")$Block, nperm = 9999) #[MA]permutations within blocks
    ) # how defines the permutations, it is important to adjust it to the experimental design such as a time series
    return(output)
  })
}

# run custom function to get results per sub-community
set.seed(5235)
perm_by4 <- permanova_with_list(ps_list_CSS, "MeJA_treatment")
perm_by4 <- lapply(perm_by4, function(perm){as.data.frame(perm)})


#[MA] get df

# [MA]Function to tidy one PERMANOVA table (add sample source as a column)
extract_perm_table <- function(perm_table, sample_name) {
  perm_table %>%
    as.data.frame() %>%
    tibble::rownames_to_column("Source") %>%
    mutate(Sample = sample_name)
}

#[MA] Apply the function to each element in the list
combined_perm <- map2_df(perm_by4, names(perm_by4), extract_perm_table)

#[MA] Split 'Sample' into Plant_species and Compartment
combined_perm <- combined_perm %>%
  separate(Sample, into = c("Plant_species", "Compartment"), sep = "\\_")

#[MA] Final combined table [This is Table S3]
print(combined_perm)

#[MA] copy to clipboard for pasting into Excel
write.table(combined_perm, "clipboard", sep = "\t", row.names = FALSE)


```

# 2.5 - Save session
```{r}

save.image("./R output/env.2_beta_diversity.RData")


normalization_listed
ps_list_CSS 


#Remove what you don't need for next script and keep phyloseq lists
rm(list = setdiff(ls(), c("normalization_listed", "ps_list_CSS")))

```

Beta Diversity analysed! proceed to scrip 3_Alpha_Diversity!

