---
title: "9_plant_traits"
author: "Marcela Aragon"
date: "8/17/2022"
editor_options: 
  chunk_output_type: console
---

# 9.0 Plant traits 

During the experiment, several plant traits were measured to characterize the plant's phenotypic response to the different stress treatments. For both *A. thaliana* and *B. oleracea* 1. Dry weight was measured. Additionally 2. Leaf area, 3. Leaf length and 4.Leaf width were measured only for *B. oleraceae* and  5.Number of fruits and 6.Inflorescence dry weight were measure only for Arabidopsis as plants were already flowering. 

Finally, as both plants are phenotypically different *per se* we compared the effects of the treatments between them using Cohen's D 

## Setting up R-studio

```{r}

# this will load or install libraries for this script
library(tidyr)
library(dplyr)
library(vegan)
library(Hmisc)
library(tibble)
library(ggpubr)
library(DHARMa)
library(fitdistrplus)
library(emmeans)
library(AICcmodavg)
library(jtools)
library(car)
library(multcompView)
library(multcomp)
library(lsr)
library(effsize)
library(gridExtra)

```

## Settings for plots

```{r}

# Setting up white background for plots
theme_set(theme_bw())

# Set colors 
#colors & treatments 
color_treatments <- c("#d7d7d7", "#f2c476", "#bd5a28", "#7aaf91") #"Control", "MeJA 0.1mM", "MeJA 1mM", "P. brassicae OS"
color_species <- c("#006c3a", "#309190") #at & bo, respectively
color_compartment <- c("#B28F45", "#846475") #endo & rhizo, respectively (not really used)

# Set axis looks
axis_looks <- theme(axis.text.x = element_text(colour = "black", size = 9,
                                               face = "bold", angle=0, hjust=0.5))+
  theme(axis.text.y = element_text(colour = "black", size = 9, face = "bold"))+
  theme(axis.title=element_text(size=11, face = "bold"))+
  theme(axis.title.y = element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0)))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  theme(legend.position="none")+
  theme(plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank())+
  theme(axis.line = element_line(colour = "black", 
                      size = 0.8, linetype = "solid"))

treatment_names <- c("control" = "Control", "MeJA_0.1" = "MeJA 0.1mM",
                      "MeJA_1.0" = "MeJA 1mM", "oral_secretion"="Caterpillar OS")

```



## Load data 
```{r}
# load raw data for traits df 
raw <- read.csv("./Data/MeJA_Plant_traits_raw.csv")
str(raw )

#removing non-necessary column 
raw  <- raw [-c(7:8,16:19)]
raw[c(1:6)] <- lapply(raw[c(1:6)], factor)

#clean df
traits <- raw

#summary
summary(traits)

#df for Atha
Atha <- traits %>% dplyr::filter(species == "Arabidopsis_thaliana") %>% 
                   mutate(aboveground_dw = ldry_weight + inflorescence) #make new column with overall aboveground biomass
#df for Bole
Bole <- traits %>% dplyr::filter(species == "Brassica_oleraceae")

#df list 
l_sp <- list(Atha, Bole)
names(l_sp) <- c("Arabidopsis_thaliana", "Brassica_oleraceae")

```


## Load functions
```{r}

#functions to check data distribution and run GLMs
source("./Code/Functions/data_distribution_and_GLM_functions.R")

```


## BO_Dry weight

### Plots

```{r}

#Boxplot Bole
p.dw.Bole <- ggplot(data = Bole,
                    mapping = aes(x = treatment, y = ldry_weight), colour= treatment)+
  geom_boxplot(mapping = aes(fill= treatment), position = position_dodge(width = 0.8), lwd=0.8)+
  geom_jitter(mapping = aes(fill = treatment),
                         shape = 21, size = 1.8, color = "black", position = position_dodge(width = 0.8))+
  scale_fill_manual(values = color_treatments)+
  labs(y="Leaves dry weight (g)", x=" ") +
  labs(title = "Bole")+
  #scale_y_continuous(expand = c(0, 0), #start graph at y = min and x = min
  #                   limits = c(0, max(Bole$ldry_weight*1.1))) +
  scale_x_discrete(labels=treatment_names)+
  ylim(0,0.6)+
  axis_looks+
  theme(legend.position = "top")

p.dw.Bole

#save plot
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/09_plant_traits",
  scale = 2.5,
  width = 500,
  height = 500,
  units = "px",
  dpi = 300),
x=c("BO_Leaf_dry_weight.svg","BO_Leaf_dry_weight.png"))




```

### Data analysis

```{r}
#Brassica 
continuous_distribution_plot_table(Bole$ldry_weight) #normal 
hist(Bole$ldry_weight)

#GLM
glm.Bole <- glm_gaussian_Withdf(l_sp$Brassica_oleraceae) #Optimal
Anova(glm.Bole$Full)
summ(glm.Bole$Full, digits = 4) #significant difference 

simulateResiduals(glm.Bole$Full, plot = T) #ok, choosing this one

#### Anova 
#LM, trying with just LM
glm.Bole2 <- lm(
  ldry_weight ~ treatment + Block, 
  data = l_sp$Brassica_oleraceae)

simulateResiduals(glm.Bole2, plot = T) #ok

anova(glm.Bole2)
summary(glm.Bole2)$r.squared

summ(glm.Bole$Optimal)
summary_model(glm.Bole$Optimal, glm.Bole$Null)

#### Anova 

#Keeping GLM
df_dw_Bo <- summary_model(glm.Bole$Full, glm.Bole$Null)

df_dw_Bo <- df_dw_Bo %>% 
            mutate(trait = "dry_weight",
                   species = "Brassica_oleraceae",
                   model="GLM",
                   distribution="Gaussian")

#Post-hoc
emm.Bole <- emmeans(ref_grid(glm.Bole$Full), pairwise ~ treatment)
plot(emm.Bole$emmeans, comparisons = TRUE)

cld <- cld(emm.Bole,
           alpha = 0.05,
           Letters = letters,
           adjust = "tukey")
cld

```

* Significant effect of treatment (p=0.004578) and no effect of Block (p=0.065722) in B. oleracea biomass (LM, p=0.0029)

--> There's only difference between 0.1 and 1% MeJA, but not against control for *B. oleraceae*, for *A. thaliana* there's no difference in dry weight. 

## BO_Leaf area

### Plots

```{r}

#Boxplot Bole
p.la.Bole <- ggplot(data = Bole,
                    mapping = aes(x = treatment, y = leaf_area), colour= treatment)+
  geom_boxplot(mapping = aes(fill= treatment), position = position_dodge(width = 0.8), lwd=0.8)+
  geom_jitter(mapping = aes(fill = treatment),
                         shape = 21, size = 1.8, color = "black", position = position_dodge(width = 0.8))+
  scale_fill_manual(values = color_treatments)+
  labs(y="Leaf area (cm2)", x=" ") +
  labs(title = "Bole")+
  #scale_y_continuous(expand = c(0, 0), #start graph at y = min and x = min
                     #limits = c(0, max(Bole$leaf_area*1.1))) +
  scale_x_discrete(labels=treatment_names)+
  ylim(0,100)+
  axis_looks+
  theme(legend.position = "top")

p.la.Bole 

#save plot
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/09_plant_traits",
  scale = 2.5,
  width = 500,
  height = 500,
  units = "px",
  dpi = 300),
x=c("BO_Leaf_area.svg","BO_Leaf_area.png"))

```

### Data analysis

```{r}

#Brassica 
continuous_distribution_plot_table(Bole$leaf_area) #normal 
hist(Bole$leaf_area)

#LM
lm.la.Bole <- lm(
  leaf_area ~ treatment + Block, 
  data = l_sp$Brassica_oleraceae)

simulateResiduals(lm.la.Bole, plot = T) #ok

Anova(lm.la.Bole)
summary(lm.la.Bole)

#Post-hoc
emm.la.Bole <- emmeans(ref_grid(lm.la.Bole), pairwise ~ treatment)

plot(emm.la.Bole$emmeans, comparisons = TRUE)

cld.la <- cld(emm.la.Bole,
           alpha = 0.05,
           Letters = letters,
           adjust = "tukey")

cld.la

```

MeJA 1% had significantly lower leaf area than the rest of the treatments. 


## AT_Number of Siliques

### Plots
```{r}

#Boxplot Atha
p.nf.Atha <- ggplot(data = l_sp$Arabidopsis_thaliana,
                    mapping = aes(x = treatment, y =number_fruits ), colour= treatment)+
  geom_boxplot(mapping = aes(fill= treatment), position = position_dodge(width = 0.8), lwd=0.8)+
  geom_jitter(mapping = aes(fill = treatment),
                         shape = 21, size = 1.8, color = "black", position = position_dodge(width = 0.8))+
  scale_fill_manual(values = color_treatments)+
  labs(y="Number of siliques / plant", x=" ") +
  labs(title = "Atha")+
  #scale_y_continuous(expand = c(0, 0), #start graph at y = min and x = min
                     #limits = c(0, max(Bole$leaf_area*1.1))) +
  scale_x_discrete(labels=treatment_names)+
  axis_looks+
  ylim(0,90)+
  theme(legend.position = "top")

p.nf.Atha
  
#save plot
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/09_plant_traits",
  scale = 2.5,
  width = 500,
  height = 500,
  units = "px",
  dpi = 300),
x=c("AT_number_siliques.svg","AT_number_siliques.png"))

```

### Data analysis
```{r}

#Arabidopsis 

#check distribution
continuous_distribution_plot_table(Atha$number_fruits) #normal - gamma 
hist(Atha$number_fruits)

#LM
lm.nf.Atha <- lm(
  number_fruits ~ treatment + Block, 
  data = l_sp$Arabidopsis_thaliana)

simulateResiduals(lm.nf.Atha, plot = T) #not ok

#GLM with Gamma distribution
glm.nf.Atha <- glm (number_fruits ~ treatment + Block,
               family = Gamma (link = "log"),
               data = l_sp$Arabidopsis_thaliana)

simulateResiduals(glm.nf.Atha, plot = T) #ok

Anova(glm.nf.Atha)
summary(glm.nf.Atha)
summ(glm.nf.Atha, digits = 4)

#Post-hoc
emm.nf.Atha <- emmeans(ref_grid(glm.nf.Atha), pairwise ~ treatment, type="response")

plot(emm.nf.Atha$emmeans, comparisons = TRUE)

cld.nf <- cld(emm.nf.Atha,
           alpha = 0.05,
           Letters = letters,
           adjust = "tukey")

cld.nf 

```
There was significantly less numbers of siliques on Arabidopsis plants treated with MeJA 1% compared to the other treatments. 


## AT_Inforescence dry weight

### Plots
```{r}

#Boxplot Atha
p.inf.Atha <- ggplot(data = l_sp$Arabidopsis_thaliana,
                    mapping = aes(x = treatment, y =inflorescence ), colour= treatment)+
  geom_boxplot(mapping = aes(fill= treatment), position = position_dodge(width = 0.8), lwd=0.8)+
  geom_jitter(mapping = aes(fill = treatment),
                         shape = 21, size = 1.8, color = "black", position = position_dodge(width = 0.8))+
  scale_fill_manual(values = color_treatments)+
  labs(y="Inflorescence dry weight (g)", x=" ") +
  labs(title = "Atha")+
  #scale_y_continuous(expand = c(0, 0), #start graph at y = min and x = min
                     #limits = c(0, max(Bole$leaf_area*1.1))) +
  scale_x_discrete(labels=treatment_names)+
  ylim(0,0.25)+
  axis_looks+
  theme(legend.position = "top")

p.inf.Atha

#save plot
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/09_plant_traits",
  scale = 2.5,
  width = 500,
  height = 500,
  units = "px",
  dpi = 300),
x=c("AT_inflorescence_dry_weight.svg","AT_inflorescence_dry_weight.png"))


```

### Data analysis

```{r}

#Arabidopsis 

#check distribution
continuous_distribution_plot_table(Atha$inflorescence) #normal - gamma 
hist(Atha$inflorescence)

#LM
lm.inf.Atha <- lm(
  inflorescence ~ treatment + Block, 
  data = l_sp$Arabidopsis_thaliana)

simulateResiduals(lm.inf.Atha, plot = T) #ok

Anova(lm.inf.Atha)
summary(lm.inf.Atha)

#Post-hoc
emm.inf.Atha <- emmeans(ref_grid(lm.inf.Atha), pairwise ~ treatment)

plot(emm.inf.Atha$emmeans, comparisons = TRUE)

cld.inf <- cld(emm.inf.Atha,
           alpha = 0.05,
           Letters = letters,
           adjust = "tukey")

cld.inf

```

Significantly less with MeJA 1% compared to control and MeJA 0.1%

#Joint Plot[Fig.S2]
```{r}

ggarrange(p.inf.Atha, p.dw.Bole, p.nf.Atha, p.la.Bole,
          nrow=2, ncol=2, common.legend = TRUE)

#save plot
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/09_plant_traits",
  scale = 2.5,
  width = 1000,
  height = 1000,
  units = "px",
  dpi = 300),
x=c("AT_BO_Traits.svg","AT_BO_Traits.png"))


```


# Cohen's D comparison 

The standardized mean difference (cohen's D) with confidence intervals were calculated for:  

* dry biomass BO
* leaf area BO
* inflorescence weight AT
* Number of siliques AT

Reference suggested by Pedro: https://nph.onlinelibrary.wiley.com/doi/full/10.1111/nph.15504 

### Cohen's D
```{r}
#making costume function using lsr package to calculate Cohen's D

cohensD_per_trait <- function(df, column, species){
  
  #df= data frame
  #column = column name with trait of interest, no quotes 
  #species = species name to be put in df, no quotes 
  
  trait_name <- deparse(substitute(column)) #gets actual name (" ") of the column to be substituted
  species_name <- deparse(substitute(species)) #idem for species

Ctrl<- df %>%  #gets only values for control and the trait 
       dplyr::filter(treatment == "control") %>%  pull(trait_name) 

MeJA_0.1 <- df %>% 
           dplyr::filter(treatment == "MeJA_0.1") %>% pull(trait_name)

MeJA_1 <- df %>% 
           dplyr::filter(treatment == "MeJA_1.0") %>% pull(trait_name)
  
OS <- df %>% 
           dplyr::filter(treatment == "oral_secretion") %>% pull(trait_name)

d_MeJA0.1 <- cohen.d(MeJA_0.1, Ctrl) #calculates Cohen's D, Ctrl should be 2nd position
d_MeJA1.0 <- cohen.d(MeJA_1, Ctrl)   #so estimate matches the effect: 
d_OS <- cohen.d(OS, Ctrl)            #(smaller or larger than ctrl)

a <- as.data.frame(t(unlist(d_MeJA0.1))) %>% #makes new df with values from Cohen's D estimation
      mutate(treatment="MeJA_0.1")           #and adds a new column with the treatment 

b <- as.data.frame(t(unlist(d_MeJA1.0))) %>% 
      mutate(treatment="MeJA_1.0")  

c <- as.data.frame(t(unlist(d_OS))) %>% 
      mutate(treatment="oral_secretion") 

df.cohensD <- rbind(a,b,c)%>%             #merge 3 df's
              mutate(trait=trait_name)%>% #makes a new column with the trait name 
              mutate(species=species_name)%>% #idem for species 
              mutate(eff_size=                #Adds a new column replacing numeric "magnitude" 
                       case_when(             #values (1-4) coming from cohen.d function 
                         magnitude == 1 ~ "negligible", #with the category written in text 
                         magnitude == 2 ~ "small",
                         magnitude == 3 ~ "medium",
                         magnitude == 4 ~ "large"))%>%
             relocate(trait, species, treatment, eff_size)


return(df.cohensD) #returns full df 
}  

#calculate cohen's D for each trait 
d_drybiomass <- cohensD_per_trait(Bole, ldry_weight, Brassica_oleraceae)
d_leaf_area <- cohensD_per_trait(Bole, leaf_area, Brassica_oleraceae)
d_number_fruits <- cohensD_per_trait(Atha, number_fruits, Arabidopsis_thaliana)
d_inflorescence <- cohensD_per_trait(Atha, inflorescence, Arabidopsis_thaliana)


#merge df's
d_traits <- rbind(d_drybiomass, d_leaf_area, d_number_fruits,d_inflorescence)

#changing to factor and numeric 
d_traits[c(1:4)] <- lapply(d_traits[c(1:4)], as.factor)
d_traits[c(7:13)] <- lapply(d_traits[c(7:13)], as.numeric)

#order traits
d_traits$trait = factor(d_traits$trait,
                         level=c("inflorescence", "number_fruits", "ldry_weight", "leaf_area"))

```

### Plots[Fig.1a]  

Followed this: https://stackoverflow.com/questions/58657802/forest-plot-with-subgroups-in-ggplot2 

```{r}

#let's make new column to plot everything at once
d_traits <- d_traits %>% mutate(Sp = case_when(
    species == "Arabidopsis_thaliana" ~ "AT",
    species == "Brassica_oleraceae" ~ "BO",
    TRUE ~ NA_character_))  # optional safety fallback

d_traits <- d_traits %>%  mutate(Sp_Trait = paste0(Sp,".", trait)) 

# Order factors so they appear in logical order on the plot
d_traits$Sp_Trait <- factor(d_traits$Sp_Trait,
                            levels = rev(c("AT.inflorescence", "AT.number_fruits",
                                       "BO.ldry_weight", "BO.leaf_area")))

d_traits$treatment <- factor(d_traits$treatment, levels = c("MeJA_0.1", "MeJA_1.0", "oral_secretion"))
d_traits$treatment <- factor(d_traits$treatment, levels = c("oral_secretion", "MeJA_1.0", "MeJA_0.1"))


#settings for plots 
color_treatments_noControl <- c("#7aaf91", "#bd5a28", "#f2c476")
names_treatments_noControl <- c(c("MeJA_0.1" = "0.1 mM \n MeJA", "MeJA_1.0" = "1 mM \n MeJA", "oral_secretion"="P. brassicae \n OS"))

#plot
p.cohensD <- ggplot(data = d_traits,
                         mapping = aes(x = Sp_Trait, y = estimate,
                    ymin=conf.int.lower, ymax=conf.int.upper,colour= treatment, fill=treatment))+
  geom_linerange(size=2.5, position=position_dodge(width = 0.6)) +
  geom_hline(aes(yintercept=0)) +
    geom_hline(aes(yintercept=0.8), lty=2, colour="gray") +
  geom_hline(aes(yintercept=-0.8), lty=2, colour="gray") +
  geom_point(size=2, colour="black", stroke = 0.5,
             position=position_dodge(width = 0.6)) +
  scale_fill_manual(values=color_treatments_noControl)+
  scale_color_manual(values=color_treatments_noControl)+
  scale_y_continuous(name="Cohen's D estimate with 95% confidence interval",
                   limits = c(-4,2))+
  coord_flip()+
  labs(title = "Cohen's D")+
  axis_looks+
  theme(axis.text.y = element_text(colour = "black", size = 12,
                    face = "bold", angle=0, hjust=0.5))+
  theme(legend.position="top")

p.cohensD 

#save plot [Figure 1a]
mapply(function(x)
  ggsave(
  plot = last_plot(),
  filename =x,
  path = "./Plots/09_plant_traits",
  #scale = 2.5,
  width = 12,
  height = 18,
  units = "cm",
  dpi = 300),
x=c("AT_BO_CohenD.svg","AT_BO_CohenD.png"))




```

#Save data
```{r}

#[MA] Let's save the environment 
save.image("./R output/env.9_plant_traits.RData")

#[MA] Now let's clean what we don't need for next script 7-Random Forest
rm(list = ls())

```





